---
title: "car_analysis"
author: "Sanjay Singh, PhD"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Study on defensive response of Chickpea genotypes against gram pod borer

Two diffrential responsive genotype against gram pod borer.
Two time points-- 0 and 24 (or, say control and treatment)

```{r cars, message=FALSE, warning=FALSE}
library(readxl)
library(DESeq2)
```

```{r}
count_data <- read_excel("C:/Users/IASRI/Dropbox/IASRI/car/count_data.xlsx", sheet = "final_inp", col_types = c("text", "skip", "skip", "skip", "skip", "skip", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric"))

metadata <- read_excel("C:/Users/IASRI/Dropbox/IASRI/car/count_data.xlsx", sheet = "metadata")
```


data summary:

```{r}
car_count <- as.data.frame(count_data[, 2:13])
rownames(car_count) <- count_data$Geneid
head(car_count)
```

```{r}
car_meta <- as.data.frame(metadata[, 2:4])
rownames(car_meta) <- metadata$sample_id
car_meta
```

```{r}
all(rownames(car_meta) %in% colnames(car_count))
```
Raw count of gene expressed:

```{r}
dds <- DESeqDataSetFromMatrix(countData = car_count,
                              colData = car_meta,
                              design = ~ condition)
dds
```
Count of gene which are more than >=10 in dataset, that are used in DE analysis:

```{r}
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
dds
```
```{r}
dds <- DESeq(dds)
ntd <- normTransform(dds)
vsd <- vst(dds, blind=FALSE)

```

Heatmap of Top 20 gene expressed- 

```{r}
library("pheatmap")
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
df <- as.data.frame(colData(dds)[,c("condition", "genotype")])
pheatmap(assay(ntd)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df)
```

Sample quality heatmap-

```{r}
sampleDists <- dist(t(assay(vsd)))
library("RColorBrewer")
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(vsd$condition, vsd$hpi, sep="-")
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors)
```

Sample PCA plot for quality; its seems form quality check plot sampling was not proper so the diffrential expression analysis will have an impact. We will get few diffrentially expressed gene.

```{r}
library(ggplot2)
pcaData <- plotPCA(vsd, intgroup=c("condition", "hpi"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=condition, shape=as.factor(hpi))) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()
```

Summary of susceptible genotype diffrential expression-

```{r}
sus_result <- results(dds, contrast=c("condition","st","sc"))
sus_result_Ordered <- sus_result[order(sus_result$pvalue),]
summary(sus_result)
```
Summary of resistant genotype diffrential expression-

```{r}
res_result <- results(dds, contrast=c("condition","rt","rc"))
res_result_Ordered <- res_result[order(res_result$pvalue),]
summary(res_result)
```
No of DEG with <0.05 FDR in resistant genotype-

```{r}
sum(res_result$padj < 0.05, na.rm=TRUE)
```
No of DEG with <0.05 FDR in susceptible genotype-

```{r}
sum(sus_result$padj < 0.05, na.rm=TRUE)
```

MA plot of resistant
```{r}
plotMA(sus_result, ylim=c(-2,2))
```

MA plot of susceptible
```{r}
plotMA(res_result, ylim=c(-2,2))
```

Result summary of resistant DEG given in excel with FDR <0.1
```{r}
res_result_Sig <- subset(res_result_Ordered, padj < 0.1)
write.csv(as.data.frame(res_result_Sig), 
          file="res_result_Sig.csv")
summary(res_result_Sig)
```
Result summary of susceptible DEG given in excel with FDR <0.1

```{r}
sus_result_Sig <- subset(sus_result_Ordered, padj < 0.1)
write.csv(as.data.frame(sus_result_Sig), 
          file="sus_result_Sig.csv")
summary(sus_result_Sig)
```
volcano plot of resistant with FDR <0.05

```{r}
res_result_v <- as.data.frame(res_result)
ggplot(res_result_v) +
  geom_point(aes(x = log2FoldChange, y = -log10(padj),
                 color = padj<0.05)) +
  xlab("log2 fold change") +
  ylab("-log10 padj") +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25)))
```

volcano plot of susceptible with FDR <0.05

```{r}
sus_result_v <- as.data.frame(sus_result)
ggplot(sus_result_v) +
  geom_point(aes(x = log2FoldChange, y = -log10(padj),
                 color = padj<0.05)) +
  xlab("log2 fold change") +
  ylab("-log10 padj") +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25)))
```

volcano plot of resistant with pvalue <0.05

```{r}
res_result_v <- as.data.frame(res_result)
ggplot(res_result_v) +
  geom_point(aes(x = log2FoldChange, y = -log10(pvalue),
                 color = pvalue<0.05)) +
  xlab("log2 fold change") +
  ylab("-log10 pvalue") +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25)))
```

volcano plot of susceptible with pvalue <0.05

```{r}
sus_result_v <- as.data.frame(sus_result)
ggplot(sus_result_v) +
  geom_point(aes(x = log2FoldChange, y = -log10(pvalue),
                 color = pvalue<0.05)) +
  xlab("log2 fold change") +
  ylab("-log10 pvalue") +
  theme(legend.position = "none",
        plot.title = element_text(size = rel(1.5), hjust = 0.5),
        axis.title = element_text(size = rel(1.25)))
```














